---
title: "Read Individual data and create basic variables"
format: html
editor: visual
---

####### This script reads the Individual dataset from the DHS. It renames a few variables for ease of use later

# load-packages

```{r, message = F, warning = F}
# rm(list = ls())
library(haven)
library(tidyverse)
library(summarytools)
```

# Read IR Data

## select variables of interest

### Meta variables
```{r IR variables of interest}
varlist_meta <- c("v021", "v025", "v023", "sdist", "v024", "v008a", "v040") 

## v021:=PSU, v025:=Rural/Urban, v023:=Strata, sdist:=District, v024:=State, 
## v008a:=Date of Interview(cdc), v040:= clust altitude in mtrs, 
```

### SES variables - Social
```{r}
varlist_ses_social <- c("v130", "v131", "s116", "v151", "v152")

## v130:= religion, v131:= ethnicity, s116:= caste, v151:=sex of HH head, 
## v152:=age of HH head
```


### SES variables - Econ
#### Household Ameneties
```{r}
varlist_ses_econ_sli <- c("v113", "v115", "v116", "v119", "v120", "v121", "v122", 
             "v123", "v124", "v125", "v127", "v128", "v129", "v153", "v161", "v190", 
             "v191", "v191a")

## v113: source of drinking water, v115: time to get to water source, 
## v116:= type of toilet facility, v119: HH has electricity, v120: HH has radio,
## v121:= HH has TV, v122:= HH has fridge, v123:= HH has bicycle, v124:=motorcycle/scooter,
## v125:= car/truck, v127:= main floor material, v128:=main wall material, v129:= roof material
## v153:=HH has telephone(land-line), v161:= type of cooking fuel, v190:= wealth index combined,
## v191:=wealth index factor score comb (5 decimals), v191a:=wealth index factor score for rural/urban
```

#### Other economic indicators
```{r}
varlist_ses_econ_other <- c("v106", "v107", "v133", "v149")

## v106:= highest edu level, v107:= highest year of edu, v133:= edu in single yrs,
## v149:= edu attainment
```


### SES variables - demographic
#### Age and Parity related
```{r}
varlist_ses_demog_gen <- c("v012", "v020", "v136", "v137", "v138", "v201", "v202", 
                       "v203", "v204", "v205", "v206", "v207", "v208", "v209", "v212")

## v012:=Respondent age, v020:=ever-married sample, v136:=no of HH members, v137:= no child under 5,
## v138:= no eligible women in HH, v201:=tot child ever born, v202:=sons at home, v203:=daughters at home,
## v204:=sons elsewhere, v205:=daughters elsewhere, v206:=sons who have died, v207:=daughters who have died, 
## v208:=births in last 5 years, v209:=births in past year, v212: age at first birth
```

### Read individual dataset

```{r read-IR}
df_dhs_IR_raw <- read_dta("../../common-datasets/india-nfhs-5/individual-recode/IAIR7DFL.DTA",
                      col_select = all_of(varlist))

df_dhs_IR_raw <- as_factor(df_dhs_IR_raw) 
```

## Rename Variables

```{r}
df_dhs_IR <- df_dhs_IR_raw %>% 
  # Renaming SES variables
  rename(psu = v021, 
         rural = v025,
         strata = v023,
         dist_name = sdist,
         state_name = v024,
         wt_raw = v005,
         psu_alt_mtrs = v040,
         mat_age = v012,
         mat_parity = v201,
         mat_edu_level = v106,
         hh_religion = v130,
         hh_caste = s116,
         hh_wealth_quintile_og = v190,
         hh_wealth_score_og = v191) %>% 
  # Create a few basic variables
  ## wealth score
  mutate(hh_wealth_score_og = hh_wealth_score_og/100000) %>% 
  # Weight
  mutate(wt = wt_raw/1000000) %>% 
  # religion - non-hindu
  mutate(hh_religion_club = case_when(
    hh_religion == "hindu" ~ "hindu",
    hh_religion == "muslim" ~ "muslim",
    hh_religion == "christian" ~ "christian",
    TRUE ~ "other")) %>% 
  mutate(hh_religion_club = fct_relevel(hh_religion_club, "hindu", "muslim", 
                                        "christian", "other")) 
  

names(df_dhs_IR)
```

## Drop respondents who are not de jure residents

```{r}
sum(is.na(df_dhs_IR$hh_caste))

## 18312 respondents will be dropped in the next step
```

```{r}

df_dhs_IR_comp_cases <- df_dhs_IR %>% 
  filter(v119 != "not a dejure resident") %>%  # 18312 cases dropped
  filter(!is.na(hh_caste)) # 34204 cases dropped
  
nrow(df_dhs_IR) - nrow(df_dhs_IR_comp_cases)
```

```{r}
df_dhs_IR_comp_cases %>% map_dbl(~sum(is.na(.)))
```
