---
title: "Read Individual data and create basic variables"
format: html
editor: source
---

####### This script reads the Individual dataset from the DHS. It renames a few variables for ease of use later

# load-packages

```{r, message = F, warning = F}
# rm(list = ls())
library(haven)
library(tidyverse)
library(summarytools)
```

# Read IR Data

## select variables of interest

### Meta variables
```{r IR variables of interest}
varlist_meta <- c("v021", "v025", "v023", "sdist", "v024", "v008a", "v040",
                  "v005", "v040") 

## v021:=PSU, v025:=Rural/Urban, v023:=Strata, sdist:=District, v024:=State, 
## v008a:=Date of Interview(cdc), v040:= clust altitude in mtrs, 
## v005 := sample weights
```

### SES variables - Social
```{r}
varlist_ses_social <- c("v130", "v131", "s116", "v151", "v152")

## v130:= religion, v131:= ethnicity, s116:= caste, v151:=sex of HH head, 
## v152:=age of HH head
```


### SES variables - Econ
#### Household Ameneties
```{r}
varlist_ses_econ_sli <- c("v113", "v115", "v116", "v119", "v120", "v121", "v122", 
             "v123", "v124", "v125", "v127", "v128", "v129", "v153", "v161", "v190", 
             "v191", "v190a", "v191a")

## v113: source of drinking water, v115: time to get to water source, 
## v116:= type of toilet facility, v119: HH has electricity, v120: HH has radio,
## v121:= HH has TV, v122:= HH has fridge, v123:= HH has bicycle, v124:=motorcycle/scooter,
## v125:= car/truck, v127:= main floor material, v128:=main wall material, v129:= roof material
## v153:=HH has telephone(land-line), v161:= type of cooking fuel, v190:= wealth index combined,
## v191:=wealth index factor score comb (5 decimals), v191a:=wealth index factor score for rural/urban
```

#### Other economic indicators
```{r}
varlist_ses_econ_other <- c("v106", "v107", "v133", "v149")

## v106:= highest edu level, v107:= highest year of edu, v133:= edu in single yrs,
## v149:= edu attainment
```


### SES variables - demographic
#### Age and Parity related
```{r}
varlist_ses_demog_gen <- c("v012", "v020", "v136", "v137", "v138", "v201", "v202", 
                       "v203", "v204", "v205", "v206", "v207", "v208", "v209", "v212")

## v012:=Respondent age, v020:=ever-married sample, v136:=no of HH members, v137:= no child under 5,
## v138:= no eligible women in HH, v201:=tot child ever born, v202:=sons at home, v203:=daughters at home,
## v204:=sons elsewhere, v205:=daughters elsewhere, v206:=sons who have died, v207:=daughters who have died, 
## v208:=births in last 5 years, v209:=births in past year, v212: age at first birth
```

### Maternal healthcare utilization
```{r}
varlist_mat_health_utiliz <- c("midx_1", "midx_2", "midx_3", "midx_4", "midx_5",
                               "m14_1", "m14_2", "m14_3", "m14_4", "m14_5", 
                               "m15_1", "m15_2", "m15_3", "m15_4", "m15_5")
```

### Empowerment Indicators
```{r}
varlist_empowerment <- c("s309", "s929", "s930a", "s930b", "s930c", "s931", "s932", 
                         "v731", "v732", "v739", "v743a", "v743b", "v743c", "v743d", 
                         "v743e", "v743f", "v745a", "v745b", "v169a", "v170")
```

### Violence Indicators
#### Exposure to IPV
```{r}
varlist_ipv_expsoure <- c(
                  c(paste0("d101",letters[1:6])), 
                    c(paste0("d103",letters[1:3])), 
                     c(paste0("d105",letters[1:14])),
                      c(paste0("d105",letters[1:14])),
                       "d106", "d107", "d108")
```

#### Justification of violence
```{r}
varlist_ipv_justified <- c("s943f", "s943g", "s943f", "s943g", 
                           c(paste0("v744",letters[1:5])))
```

### Male participation
```{r}
varlist_male_participation <- c("s422_1", "s422_2", "s422_3", "s422_4", "s422_5", "s422_6",
                                "s909", "s910", "s920", "s301")
```

### Frequency of using media
```{r}
varlist_media <- c("v157", "v158", "v159")
```

### Accessibility to maternal healthcare
```{r}
varlist_healthcare_access <- c(
                              c(paste0("m65a_",c(1:6))),
                               c(paste0("m65b_",c(1:6))),
                                c(paste0("m65c_",c(1:6))),
                                 c(paste0("m65d_",c(1:6))),
                                  c(paste0("m65e_",c(1:6))),
                                   c(paste0("m65f_",c(1:6))),
                                    c(paste0("m65g_",c(1:6))),
                                     c(paste0("m65h_",c(1:6))),
                                      c(paste0("v467",letters[1:9])), 
                                               "v481", "s359", "s361") 

```


## Varlist merged
```{r}
varlist_select <- c(varlist_meta, varlist_ses_social, varlist_ses_econ_sli, 
                       varlist_ses_econ_other, varlist_ses_demog_gen, 
                    varlist_mat_health_utiliz, varlist_empowerment, 
                    varlist_ipv_expsoure, varlist_ipv_justified, 
                    varlist_male_participation, varlist_media, 
                    varlist_healthcare_access)
```

### Read individual dataset

```{r read-IR}
df_dhs_IR_raw <- read_dta("../../common-datasets/india-nfhs-5/individual-recode/IAIR7DFL.DTA",
                      col_select = all_of(varlist_select))

df_dhs_IR_raw <- as_factor(df_dhs_IR_raw) 
df_dhs_IR_raw_dt <- setDT(df_dhs_IR_raw)
beep(sound = 5)
```

## Identify variables that have missing cases
```{r}
df_missing <- df_dhs_IR_raw_dt %>% map_dbl(~sum(is.na(.))) 
```

## Rename Variables

```{r}
df_dhs_IR <- setnames(df_dhs_IR_raw_dt,
                      old = c("v021", "v025", "v023", "sdist", "v024",
                              "v005", "v040", "v012", "v201", "v106", "v130",
                              "s116", "v190", "v191", "v190a"),
                      new = c("psu", "rural", "strata", "dist_name", "state_name", 
                              "wt_raw", "psu_alt_mts", "mat_age", "mat_parity", 
                              "mat_edu_level", "hh_religion", "hh_caste",
                              "hh_wealth_quintile_og", "hh_wealth_score_og", 
                              "hh_wealth_quintile_ru_og"))
```


```{r}
freq(df_dhs_IR$hh_caste)
```


## Drop cases

```{r}
df_dhs_IR <- df_dhs_IR[
  # Drop cases if caste is missing or dont know
  hh_caste != "don't know" & !is.na(hh_caste)] # 38,691 cases dropped
```



